### 2016-10-07 dissertation/normal.tex
dissertation chapter Normal: update 4

\begin{figure}[H]
\centering{
\includegraphics[width=3.5in]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph}
\label{fig:Moore}}
\end{figure}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=3.5in]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
\label{fig:convention}}
\end{figure}

@@

The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
However, what makes NB powerful is its property when doing multiplications and exponentiations.
The following lemmas and examples illustrate this fabulous property very well.
\begin{Lemma}[Square of NB]
In $\Fkk$, equation 
\begin{equation*}
(a+b)^2 = a^2 + b^2
\end{equation*}
has been proved. According to the \textbf{binomial theorem}, it can be extended to
\begin{align*}
&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{k-1}})^2 \\
&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
&= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
\end{align*}
\end{Lemma}
This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
Obviously, StdB representation does not have this benefit.

@@

\begin{Example}
\emph{NB multiplication:} Assume there are 2 binary vectors representing 2 operands in normal
basis: $A = (a_0, a_1, \dots, a_{n-1}), B = (b_0, b_1, \dots, b_{n-1})$; similarly the product can also be written
as: $C = A*B = (c_0, c_1, \dots, c_{n-1}).$

Then the highest digit of product can be represented by a function: $c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, 
\dots, b_{n-1})$. Square both side: $C^2 = A^2*B^2$, i.e. the second highest digit $c_{n-2} = f(a_{n-1}, a_0, a_1, 
\dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2}).$ By this method it is easy to get all digits of product $C$.
\end{Example}

>>>>

\begin{figure}[H]
\centering{
\includegraphics[width=5in]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph}
\label{fig:Moore}}
\end{figure}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=5in]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
\label{fig:convention}}
\end{figure}

@@

The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
However, what makes NB powerful is its property when doing multiplications and exponentiations.
The following lemmas and examples illustrate this fabulous property very well.
\begin{Lemma}[Square of NB]
\label{lem:squareNB}
In $\Fkk$, equation 
\begin{equation*}
(a+b)^2 = a^2 + b^2
\end{equation*}
has been proved. According to the \textbf{binomial theorem}, it can be extended to
\begin{align*}
&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
&= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
\end{align*}
\end{Lemma}
This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
Obviously, StdB representation does not have this benefit.

@@

This example shows that convenience to use NB when computing $2^k$ power of an element.
Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
operations, the property in lemma \ref{lem:squareNB} can be well utilized.

\begin{Example}[Bit-wise NB multiplication]
Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
$A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$

Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
\begin{equation}
\label{eqn:multiMSB}
c_{n-1} = f_{mult}(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})
\end{equation}
Before discussing the details of the function $f_{mult}$, we can take 
a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
becomes the new MSB because of right-cyclic shifting. 
Concretely, 
$$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
$$c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})$$ 
Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
\end{Example}

From above example, it is proved that a universal structure that implements $f_{mult}$ can be reused
for $k$ times in NB multiplication over $\Fkk$. Comparing to StdB, which requires distinct design 
for every bit of multiplication, NB is less costly if we can prove $f_{mult}$ will not result in 
a structure with $O(k^2)$ complexity. So our next mission is to explore the details of $f_{mult}$
to prove it will be a relatively simple design with complexity lower than $O(k^2)$.

If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
linear functions. As we know, matrix multiplication can simulate all possible combinations of 
linear functions (which is also the reason it is used as basic model to simulate the behavior of a neuron
in neural network machine learning algorithms). Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
column vector, then the single bit product can be written as the product of matrix multiplication
$$c_{i} = A\times C\times B$$
where $C$ is a $k\times k$ square matrix.
